---
title: "Getting Started With preText"
author: "Matthew Denny"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting_started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette is designed to introduce you to the **preText** R package. This 
package is built on top of the [quanteda](https://github.com/kbenoit/quanteda)
R package for text processing and can take as input a `quanteda::corpus` object,
or a character vector (with one string per document). The main functions will
preprocess the input text 64-128 different ways, and then allow the user to 
assess the consequences of the different preprocessing steps based on how 
unusual the resulting document--term matrix (DTM) is using the **preText**
procedure.

Our paper detailing the **preText** procedure can be found at the link below:

* Matthew J. Denny, and Arthur Spirling (2016). "Assessing the Consequences of Text Preprocessing Decisions". [[ssrn.com/abstract=2849145]](https://ssrn.com/abstract=2849145)

## Installation
The release version of the package can be installed from CRAN as follows:
```{r eval=FALSE} 
install.packages("preText")
```  

If you want to get the latest version from GitHub, start by checking out the 
**Requirements for using C++ code with R** section in the following 
tutorial: [Using C++ and R code Together with Rcpp](http://www.mjdenny.com/Rcpp_Intro.html). 
You will likely need to install either `Xcode` or `Rtools` depending on whether 
you are using a Mac or Windows machine before you can install the preText package 
via GitHub, since it makes use of C++ code.

```{r eval=FALSE} 
install.packages("devtools")
```    
Now we can install from GitHub using the following line:
```{r eval=FALSE} 
devtools::install_github("matthewjdenny/preText")
```
Once the `GERGM` package is installed, you may access its functionality as you 
would any other package by calling:
```{r eval=FALSE} 
library(preText)
```
If all went well, check out the `vignette("getting_started_with_preText")` which 
will pull up this vignette!

## Basic Usage
We begin by loading the package and some example data from the `quanteda` R 
package. In this example, we will make use of 57 U.S. presidential inaugural 
speeches. As a general rule, you will want to limit the number of documents used
with `preText` to several hundred in most cases, in order to avoid extremely 
long run times and/or high memory requirements. To make this example run more 
quickly, we are only going to use 10 documents.
```{r eval=TRUE, fig.width=6, fig.height=6, fig.align ='center'}     
library(preText)
library(quanteda)

# load in U.S. presidential inaugural speeches from Quanteda example data.
corp <- corpus(inaugTexts)
# use first 10 documents for example
documents <- texts(corp)[1:10]
# take a look at the document names
print(names(documents))
```
Having loaded in some data, we can now make use of the `factorial_preprocessing()` 
function, which will preprocess the data 64 or 128 different ways (depending on
whether n-grams are included). In this example, we are going to preprocess the 
documents all 128 different ways. This should take between 5 and 10 minutes on 
most modern laptops. Longer documents and larger numbers of documents will 
significantly increase run time and memory usage. It is highly inadvisable to use 
more than 500-1,000 under any circumstances and in the case where the user wishes 
to preprocess more than a few hundred documents, they may want to explore the 
`parallel` option. This can significantly speed up preprocessing, but will 
require significantly more RAM on the computer being used. Here, we have 
selected the `use_ngrams = TRUE` option, and set the document proportion 
threshold at which to remove infrequent terms at 0.2. This means that terms 
which appear in less than 20 percent (2/10) documents will be removed. The 
default value is 0.01 (or 1/100 documents), but for this small corpus, we 
increase the value. In order prevent spamming this vignette with output, we have 
elected to set the  `verbose` option to FALSE. In practice, it is better to keep 
`verbose = TRUE` to make it easier to evaluate the progress of preprocessing.
```{r eval=TRUE, fig.width=6, fig.height=6, fig.align ='center'}
preprocessed_documents <- factorial_preprocessing(
    documents,
    use_ngrams = TRUE,
    infrequent_term_threshold = 0.2,
    verbose = FALSE)
```
 This function will output a list object with 
three fields. The first of these is `$choices`, a data.frame containing indicators
for each of the preprocessing steps used. The second is `$dfm_list`, which is a 
list with 64 or 128 entries, each of which contains a `quanteda::dfm` object 
preprocessed according to the specification in the corresponding row in `choices`.
Each DFM in this list will be labeled to match the row names in choices, but you
can also access these labels from the `$labels` field. We can look at the first 
few rows of `choices` below:
```{r eval=TRUE, fig.width=6, fig.height=6, fig.align ='center'}
names(preprocessed_documents)
head(preprocessed_documents$choices)
```
Now that we have our preprocessed documents, we can perform the preText 
procedure on the factorial preprocessed corpus using the `preText()` function.
It will be useful now to give a name to our data set using the `dataset_name`
argument, as this will show up in some of the plots we generate with the output.
The standard number of pairs to compare is 50 for reasonably sized corpora, but
because we are only using 10 documents, the maximum number of pairwise document
distances is only (10)*(10 - 1)/2 = 45, so we select 20 pairwise comparisons for
purposes of illustration. This function will usually not take as long to run as 
the `factorial_preprocessing()` function, but parallelization is also available 
for this function if a speedup is desired. It is suggested that the user select
`verbose = TRUE` in practice, but we set it to FALSE here to avoid cluttering 
this vignette. This function should run in 10-30 seconds for this small corpora, 
and in several hours to a day for most moderately sized corpora.
```{r eval=TRUE, fig.width=6, fig.height=6, fig.align ='center'}
preText_results <- preText(
    preprocessed_documents,
    dataset_name = "Inaugural Speeches",
    distance_method = "cosine",
    num_comparisons = 20,
    verbose = FALSE)
```
The `preText()` function returns a list of result with four fields:

* `$preText_scores`: A `data.frame` containing preText scores and preprocessing step labels for each preprocessing step as columns. Note that there is no preText score for the case of no prepprocessing steps. 
* `$ranked_preText_scores`: A `data.frame` that is identical to `$preText_scores` except that it is ordered by the magnitude of the preText score
* `$choices`: A `data.frame` containing binary indicators of which preprocessing 
steps were applied to factorial preprocessed DFM. 
* `$regression_results`: A `data.frame` containing regression results where indicators for each preprocessing decision are regressed on the preText score for that specification.

We can now feed these results to two functions that will help us make better 
sense of them. `preText_score_plot()` creates a dot plot of scores for each 
preprocessing specification:
```{r eval=TRUE, fig.width=6, fig.height=16, fig.align ='center'}
preText_score_plot(preText_results)
```
Here, the least risky specifications have the lowest preText score and are 
displayed at the top of the plot. We can also see the conditional effects of 
each preprocessing step on the mean preText score for each specification that 
included that step. Here again, a negative coefficient indicates that a step 
tends to reduce the unusualness of the results, while a positive coefficient
indicates that applying the step is likely to produce more unusual results for 
that corpus.
```{r eval=TRUE, fig.width=6, fig.height=4, fig.align ='center'}
regression_coefficient_plot(preText_results,
                            remove_intercept = TRUE)
```
In this particular toy example, we see that including n-grams and removing 
stop words tends to produce more normal results, while removing punctuation
tends to produce more unusual results. However, we should not put too much stock
in these results due to the small number of documents used to generate them.

## More Advanced Features
The preText package provides a number of additional functions for examining the
effects of preprocessing decisions on resulting DFMs. Please see the README on
the package [GitHub page](https://github.com/matthewjdenny/preText) for more
details on these additional functions.

## Replication of Results for UK Manifestos Data
The example provided above uses a toy dataset, mostly to reduce the runtime for 
the analysis to under 20 minutes on most computers. For those who would like
to explore a full example, we include the UK Manifestos dataset described in the 
paper with our package. It can be accessed using the `data("UK_Manifestos")`
command once the package is loaded. Below, we provide a full working example 
which will replicated results from the paper. The code has not been run in this 
vignette to save time, but you may give it a try on your own computer. It should 
run in less than 24 hours on most computers.
```{r eval=FALSE, fig.width=6, fig.height=4, fig.align ='center'}
# load the package
library(preText)
# load in the data
data("UK_Manifestos")
# preprocess data
preprocessed_documents <- factorial_preprocessing(
    UK_Manifestos,
    use_ngrams = TRUE,
    infrequent_term_threshold = 0.02,
    verbose = TRUE)
# run preText
preText_results <- preText(
    preprocessed_documents,
    dataset_name = "Inaugural Speeches",
    distance_method = "cosine",
    num_comparisons = 100,
    verbose = TRUE)
# generate preText score plot
preText_score_plot(preText_results)
# generate regression results
regression_coefficient_plot(preText_results,
                            remove_intercept = TRUE)
```

